---
title       : Influence Analysis Demonstration
subtitle    : 
author      : OmaymaS
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap,selfcontained}
mode        : standalone # {standalone, draft,selfcontained}


---

## Background

### - Influence Analysis is performed to evaluate influential observations in regression models.

### - A simple interactive Shiny App is presented here to clarify the intuition behind leverage and influence.



---  .class #id

## Application Overview

- The app demonstrates the leverage and influence of an adjustable point that is part of a dataset with 101 points; 100 of which are normally distributed. You can select the X-Y coordinates of the adjustable point using the following sliders with ranges from [-10,10] for both X and Y. The adjustable point appears in red on the graph.

- For every point you choose, The Linear regression model coefficients are updated and also some of the influence measures such as *hatvalue*, *residual* and *dfbeta* are shown.

---  .class #id

## Application Implementation

- We start by generating 100 normally distributed data points

```{r echo=F}
#generate normally distributed 100 points
        set.seed(1523)
        n <- 100
        x1 <- rnorm(n) ; x1<-x1-mean(x1) #center the data
        y1 <- rnorm(n) ; y1<-y1-mean(y1)
```

- the point 101 is added according to the user's choice by using the sliders. For example if the user chooses [10,5] it will be prepended to the previously generated points as follows:

```{r echo=F}
c<-c(10,5)
datX<-c(c[1],x1)
datY<-c(c[2],y1)
newDF<-data.frame(X=datX,Y=datY)
```
```{r}
head(newDF)
```

---  .class #id

## Influence Analysis Measures

- The app gives the user the chance to play with different points to understand the concepts of leverage and influence. A plot is also shown to see the location of the point and how the regression line moves as a lever according to the chosen point coordinates. 


```{r echo=F}
fit<-lm(Y~X,data=newDF)
plot(Y~X,data=newDF)
abline(fit)

```

---  .class #id
## App Link and Further Explanation

[Shiny App](https://omaymas.shinyapps.io/Influence_Analysis/)

[Blog Post : Leverage and Influence in a Nutshell](http://omaymas.github.io/InfluenceAnalysis/)

